### 2019-01-30 microbiome_online.Rmd
Imported Scripts

---
title: "R Notebook"
output: html_notebook
---

This is an [R Markdown](http://rmarkdown.rstudio.com) Notebook. When you execute code within the notebook, the results appear beneath the code.

Combining 2 dat asets

ps_merge <- merge_phyloseq(pseq, atlas1006)

 get_variable(batmicrobiome,""LATITUDE"" )

 get_variable(batmicrobiome,"LONGITUDE" )

sample_data(batmicrobiome)

sample_names(batmicrobiome)

batmicrobiome[LONGITUDE] <- sample(50:1000)

sample_data(atlas1006)$human = factor( get_variable(atlas1006) %in% c("Feces", "Mock", "Skin", "Tongue") )

sample_data(atlas1006)$human = sample(50:1000)

sample_data(atlas1006)$human = rep(3,1151) %%% THIS WORKS !!!!

sample_data(atlas1006)$LATITUDE = rep(17,1151)
sample_data(atlas1006)$LONGITUDE = rep(-77,1151) 
get_variable(atlas1006,"LATITUDE" )
map_phyloseq(atlas1006)

map_phyloseq(batmicrobiome)

rep(3,12)

devtools::install_github("UrbanInstitute/urbnmapr")

atlas1006

sample_variables(atlas1006)

sample_variables(batmicrobiome)[56]  THIS WORKS

batmicrobiome = sample_data(data.frame(
  LONGITUDE = sample(50:1000)))

    sampledata = sample_data(data.frame(
  Location = sample(LETTERS[1:4], size=nsamples(physeq), replace=TRUE),
  Depth = sample(50:1000, size=nsamples(physeq), replace=TRUE),
  row.names=sample_names(physeq),
  stringsAsFactors=FALSE

  sampledata = sample_data(data.frame(
  Location = sample(LETTERS[1:4], size=nsamples(physeq), replace=TRUE),
  Depth = sample(50:1000, size=nsamples(physeq), replace=TRUE),
  row.names=sample_names(physeq),
  stringsAsFactors=FALSE
))

Try executing this chunk by clicking the *Run* button within the chunk or by placing your cursor inside it and pressing *Cmd+Shift+Enter*.

```{r}
	setwd("/omega/bashscripts")

	## Get data
	library(foreign)
	library(stargazer)
	library(psych)
	library(plyr)
	library(xtable)
	library(car)
	#library(dplyr)
	library(scales)
	library(reshape) 
	library(gmodels)
	library(nnet)
	library(tidyr)
	library(sjmisc)
	library(RColorBrewer)
  library(htmltools)
	library(broom)

library(hrbrthemes)
library(gcookbook)
library(tidyverse)

library(phyloseq)
library(microbiome)
library(knitr)
library(ggpubr)

library(phylogeo)

data(atlas1006)   
# Rename the example data (which is a phyloseq object)
pseq <- atlas1006

```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

```{r}
tab <- global(pseq, index = "all")
stargazer(tab, type = "text")

```

```{r}
####### Alpha diversity #######
ps1 <- prune_taxa(taxa_sums(pseq) > 0, pseq)
tab <- diversities(ps1, index = "all")
stargazer(tab, type = "text",summary = NULL)
kable(head(tab))
```

```{r}
#  Prepare data for vizualisation
ps1.meta <- meta(ps1)
kable(head(ps1.meta))

stargazer(ps1.meta, type = "text",summary = NULL)
# Add the diversity table to metadata
ps1.meta$Shannon <- tab$shannon 
ps1.meta$InverseSimpson <- tab$inverse_simpson

# Compare differences in Shannon index between bmi group of the study subjects

# create a list of pairwise comparisons
bmi <- levels(ps1.meta$bmi_group) # get the variables

# make a pairwise list that we want to compare.
bmi.pairs <- combn(seq_along(bmi), 2, simplify = FALSE, FUN = function(i)bmi[i])

print(bmi.pairs)
```

```{r}
# CREATING VIOLIN PLOT
# the number in the palettes must equal the number of groups in bmi.pairs
p1 <- ggviolin(ps1.meta, x = "bmi_group", y = "Shannon",
 add = "boxplot", fill = "bmi_group", palette = c("#a6cee3", "#b2df8a", "#fdbf6f","#a6cee3", "#b2df8a", "#fdbf6f")) # 
print(p1)
```

```{r}

# Pairwise comparision using non-parametric test (Wilcoxon test).
p1 <- p1 + stat_compare_means(comparisons = bmi.pairs) 
print(p1)
```

```{r}
####### RICHNESS #######

tab <- richness(pseq)
kable(head(tab))
stargazer(tab, type = "text",summary = NULL)
plot(tab)

```

```{r}
###### DOMINANCE #####

# The dominance index refers to the abundance of the most abundant species. Various dominance indices are available (see the function help for a list of options).

# Absolute abundances for the single most abundant taxa in each sample
tab <- dominance(pseq, index = "all")
kable(head(tab))
stargazer(tab, type = "text",summary = NULL)
plot(tab)

```

```{r}
######### RARENESS and low abundance ######

tab <- rarity(pseq, index = "all")
kable(head(tab))
stargazer(tab, type = "text",summary = NULL)
plot(tab)
```

```{r}
#### COVERAGE

# The coverage index gives the number of groups needed to have a given proportion of the ecosystem occupied (by default 0.5 ie 50%).

tab <- coverage(pseq, threshold = 0.5)
kable(head(tab))
stargazer(tab, type = "text",summary = NULL)
plot(tab)
```

```{r}
#### ABUNDANCE

# The core_abundance function refers to the relative proportion of the core species. Non-core abundance provides the complement (1-x; see noncore_abundance).

tab <- core_abundance(pseq, detection = .1/100, prevalence = 50/100)
kable(head(tab))
stargazer(tab, type = "text",summary = NULL)
plot(tab)

```

```{r}
##### GINI INDEX

#Gini index is a common measure for inequality in economical income. The inverse gini index (1/x) can also be used as a community diversity measure.

tab <- inequality(pseq)
kable(head(tab))
plot(tab)
```

```{r}
###### EVENNESS

# Various evenness measures are also available.

tab <- evenness(pseq, "all")
kable(head(tab))
plot(tab)
```

MICROBIOME COMPOSITION

```{r}
# Just use prevalent taxa to speed up examples
# (not absolute counts used in this example)

# Pick sample subset
library(phyloseq)

library(hrbrthemes)
library(gcookbook)
library(tidyverse)
library(dplyr)
library(microbiome)

######## THIS ALTERED THE CORE DATA ###############
#pseq <- core(atlas1006, detection = 10^3, prevalence = 95/100)
#pseq2 <- subset_samples(pseq, group == "DI" & nationality == "AFR" & timepoint.within.group == 1)

# Try another theme
# from https://github.com/hrbrmstr/hrbrthemes

# Limit the analysis on core taxa and specific sample group
p <- plot_composition(pseq,
              taxonomic.level = "OTU",
                      sample.sort = "nationality",
                      x.label = "nationality") +
     guides(fill = guide_legend(ncol = 1)) +
     scale_y_percent() +
     labs(x = "Samples", y = "Relative abundance (%)",
                                   title = "Relative abundance data",
                                   subtitle = "Subtitle",
                                   caption = "Caption text.") + 
     theme_ipsum(grid="Y")
print(p)

# Averaged by group
p <- plot_composition(pseq,
                      average_by = "bmi_group", transform = "compositional")
print(p)

```

```{r}
tmp <- plot_composition(pseq, plot.type = "heatmap", transform = "compositional", 
            sample.sort = "neatmap", otu.sort = "neatmap", mar = c(6, 13, 1, 1))

print(tmp)
```

PLOTTING TAXA Prevlence

```{r}
data(atlas1006)

# Use sample and taxa subset to speed up example
p0 <- subset_samples(atlas1006, DNA_extraction_method == "r")

# Define detection and prevalence thresholds to filter out rare taxa
p0 <- core(p0, detection = 10, prevalence = 0)

# For the available taxonomic levels
plot_taxa_prevalence(p0, "Phylum", detection = 10)
```

PLOTTING RELAVANCE ABUNDANCE

```{r}

library(phyloseq)
ps2 <- data(atlas1006)
# for example purpose we will remove samples with less than 10000
ps2 = prune_samples(sample_sums(ps1)>=2000, ps1)

# To speed up the example we will use only those OTUs that are detected 100 times and present in 50% of the samples.

pseq2 <- microbiome::core(ps2, detection = 100, prevalence = .5)

# Improve the plotting MODIFIED 
tax_table(pseq2)[tax_table(pseq2)[,"Phylum"]== "Clostridium", "Phylum" ] <- "Clostridium Phylum"

# We will also remove the "f__" patterns for cleaner labels
tax_table(pseq2)[,colnames(tax_table(pseq2))] <- gsub(tax_table(pseq2)[,colnames(tax_table(pseq2))],pattern="[a-z]__",replacement="")

# merge at family level.

pseq.fam <- aggregate_taxa(pseq2, "Phylum")

p.fam <- plot_composition(pseq.fam, sample.sort = NULL, otu.sort = NULL,
  x.label = "ibd_subtype", plot.type = "barplot", verbose = FALSE)

print(p.fam)

```

PLOTTING RELTIVE ABUNDANCE

```{r}

### REQUIREDS THE DATA FROM ABOVE 
pseq.famrel <- transform(pseq.fam, "compositional")

p.famrel <- plot_composition(pseq.famrel, sample.sort = NULL, otu.sort = NULL,
  x.label = "ibd_subtype", plot.type = "barplot", verbose = FALSE)

print(p.famrel)

# further improvements can be done as follows

p.famrel <- plot_composition(pseq.famrel, 
                             sample.sort = NULL, 
                             otu.sort = NULL, 
                             x.label = "ibd_subtype", 
                             plot.type = "barplot", 
                             verbose = FALSE) + 
  guides(fill = guide_legend(ncol = 1)) + 
  scale_y_percent() + 
  labs(x = "Samples", 
       y = "Relative abundance (%)",
       title = "Relative abundance data", 
       subtitle = "Subtitle",
       caption = "Caption text.") + 
     theme_ipsum(grid="Y")

print(p.famrel)
```

HEAT MAP

```{r}
pseq.famlog <- transform(pseq.fam, "log10")

p.famrel.heatmap <- plot_composition(pseq.famlog, 
                             sample.sort = NULL, 
                             otu.sort = NULL, 
                             x.label = "ibd_subtype", 
                             plot.type = "heatmap", 
                             verbose = FALSE)

print(p.famrel.heatmap)
```

Plot taxa prevalence

```{r}

library(microbiome)
data(peerj32)

# Rename the data
pseq <- peerj32$phyloseq

# Calculate compositional version of the data
# (relative abundances)
pseq.rel <- microbiome::transform(pseq, "compositional")

core.taxa.standard <- core_members(pseq.rel, detection = 0, prevalence = 50/100)

```

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file).

```{r}
library(ggplot2)
library(gridExtra)
library(phylogeo)

data(mountainsoil)

```

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.