### 2019-01-30 MRD_microbiome_working.Rmd
updated

---
title: "MRD_microbiome_working"
author: "Reis-Dehabadi, M"
date: "1/30/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}

library(microbiome)
library(knitr)
library(ggpubr)
library(dplyr)

# Data from http://dx.doi.org/10.1038/ncomms7342
data(atlas1006)   
# Rename the example data (which is a phyloseq object)
pseq <- atlas1006

data(dietswap)  
pseqds <- dietswap
print(dietswap)

```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}

# Richness
tab2 <- richness(pseqds)
kable(head(tab2))

```

ou can also embed plots, for example:

```{r , echo=FALSE}

# Richness
tab2 <- richness(pseqds)
kable(head(tab2))

```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.