### 2019-06-25 import_popln_data.Rmd
updated

---
title: "MRD_final_microbiome"
author: "Reis-Dehabadi, M"
date: "4/10/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '~/globalmicrobiome')

```

## Data Set-up
```{r Data set-up}

#setwd("/Users/maryam/Scripts/Microbiome")

	## Get data
	library(foreign)
	library(stargazer)
	library(psych)
	library(plyr)
	library(xtable)
	library(car)
	#library(dplyr)
	library(scales)
	library(reshape) 
	library(gmodels)
	library(nnet)
	library(tidyr)
	library(sjmisc)
	library(RColorBrewer)
  library(htmltools)
	library(broom)

  #library(hrbrthemes)
library(gcookbook)
library(tidyverse)
 #library(gdtools)

#The phyloseq package seeks to address issues with multiple microbiome analysis packages by providing a set of functions that internally manage the organizing, linking, storing, and analyzing of phylogenetic sequencing data. In general, this package is used for UniFrac analyses.
library(phyloseq)
library(microbiome)
library(knitr)
library(ggpubr)
#library(phylogeo)
library(dplyr)
#Graphing package used in phyloseq. To edit the default setting of a plot, you need to use functions in this package.
library(ggplot2)
library(viridis) 
library(cowplot)
# install.packages("devtools")
library(devtools)
# install_github("opisthokonta/tsnemicrobiota")
library(tsnemicrobiota)
# check you've installed this library
library(forcats)

# Data from http://dx.doi.org/10.1038/ncomms7342
#This data set contains genus-level microbiota profiling with HITChip for 1006 western adults with no reported health complications.

setwd("~/globalmicrobiome")
mnd.incidence <- read_csv("mnd.csv",col_names = TRUE)
mnd.incidence <- as.data.frame(mnd.incidence)

mb.gather <- function (x) {
    # Creates an empty data frame
    mb.group <- data.frame("Location" = c("Central Europe","Eastern Europe","Scandinavia", "Southern Europe","UKIE","USA"), "Mean" = c(1:6)) 
     # groups the countries to match the microbiome classification 
  c.europe <- filter(x,Location=="Belgium"|Location=="Denmark"|Location=="Germany"|Location=="The Netherlands")
    e.europe <- filter(mnd.incidence,Location=="Poland")
    scan <- filter(mnd.incidence,Location=="Finland"|Location=="Norway"|Location=="Sweden")
    s.europe <- filter(mnd.incidence,Location=="France"|Location=="Italy"|Location=="Serbia"|Location=="Spain")
    ukie <- filter(mnd.incidence,Location=="United Kingdom"|Location=="Ireland")
    usa <- filter(mnd.incidence,Location=="United States")
    # gmeans the data and put it into the data frame
    mb.group[1,2] <- mean(c.europe$Value)
    mb.group[2,2] <- mean(e.europe$Value)
    mb.group[3,2] <- mean(scan$Value)
    mb.group[4,2] <- mean(s.europe$Value)
    mb.group[5,2] <- mean(ukie$Value)
    mb.group[6,2] <- mean(usa$Value)
    return(mb.group)
    }

temp <- mb.gather(mnd.incidence)

```