### 2019-06-26 import_popln_data.Rmd
updated

---
title: "MRD_final_microbiome"
author: "Reis-Dehabadi, M"
date: "4/10/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '~/globalmicrobiome')

```

## Data Set-up
```{r Data set-up}

#setwd("/Users/maryam/Scripts/Microbiome")

	## Get data
	library(foreign)
	library(stargazer)
	library(psych)
	library(plyr)
	library(xtable)
	library(car)
	#library(dplyr)
	library(scales)
	library(reshape) 
	library(gmodels)
	library(nnet)
	library(tidyr)
	library(sjmisc)
	library(RColorBrewer)
  library(htmltools)
	library(broom)

  #library(hrbrthemes)
library(gcookbook)
library(tidyverse)
 #library(gdtools)

#The phyloseq package seeks to address issues with multiple microbiome analysis packages by providing a set of functions that internally manage the organizing, linking, storing, and analyzing of phylogenetic sequencing data. In general, this package is used for UniFrac analyses.
library(phyloseq)
library(microbiome)
library(knitr)
library(ggpubr)
#library(phylogeo)
library(dplyr)
#Graphing package used in phyloseq. To edit the default setting of a plot, you need to use functions in this package.
library(ggplot2)
library(viridis) 
library(cowplot)
# install.packages("devtools")
library(devtools)
# install_github("opisthokonta/tsnemicrobiota")
library(tsnemicrobiota)
# check you've installed this library
library(forcats)

# Data from http://dx.doi.org/10.1038/ncomms7342
#This data set contains genus-level microbiota profiling with HITChip for 1006 western adults with no reported health complications.

mb.gather <- function (x) {
    # Creates an empty data frame
    mb.group <- data.frame("Location" = c("Central Europe","Eastern Europe","Scandinavia", "Southern Europe","UKIE","USA"), "Mean" = c(1:6)) 
     # groups the countries to match the microbiome classification 
  c.europe <- filter(x,Location=="Belgium"|Location=="Denmark"|Location=="Germany"|Location=="The Netherlands")
    e.europe <- filter(x,Location=="Poland")
    scan <- filter(x,Location=="Finland"|Location=="Norway"|Location=="Sweden")
    s.europe <- filter(x,Location=="France"|Location=="Italy"|Location=="Serbia"|Location=="Spain")
    ukie <- filter(x,Location=="United Kingdom"|Location=="Ireland")
    usa <- filter(x,Location=="United States")
    # gmeans the data and put it into the data frame
    mb.group[1,2] <- mean(c.europe$Value)
    mb.group[2,2] <- mean(e.europe$Value)
    mb.group[3,2] <- mean(scan$Value)
    mb.group[4,2] <- mean(s.europe$Value)
    mb.group[5,2] <- mean(ukie$Value)
    mb.group[6,2] <- mean(usa$Value)
    return(mb.group)
    }

setwd("~/globalmicrobiome")
# files nwed to be in the owrking directory

# For MND
mnd.incidence <- as.data.frame(read_csv("./rawdata/MND-I-B.csv",col_names = TRUE))

mnd.i.12 <- as.data.frame(read_csv("./rawdata/MND-I-B-12.csv",col_names = TRUE))
mnd.i.13 <- as.data.frame(read_csv("./rawdata/MND-I-B-13.csv",col_names = TRUE))
mnd.i.14 <- as.data.frame(read_csv("./rawdata/MND-I-B-14.csv",col_names = TRUE))
mnd.i.15 <- as.data.frame(read_csv("./rawdata/MND-I-B-15.csv",col_names = TRUE))
mnd.i.16 <- as.data.frame(read_csv("./rawdata/MND-I-B-16.csv",col_names = TRUE))

mnd.prev <- as.data.frame(read_csv("./rawdata/MND-P-B.csv",col_names = TRUE))
mnd.p.12 <- as.data.frame(read_csv("./rawdata/MND-P-B-12.csv",col_names = TRUE))
mnd.p.13 <- as.data.frame(read_csv("./rawdata/MND-P-B-13.csv",col_names = TRUE))
mnd.p.14 <- as.data.frame(read_csv("./rawdata/MND-P-B-14.csv",col_names = TRUE))
mnd.p.15 <- as.data.frame(read_csv("./rawdata/MND-P-B-15.csv",col_names = TRUE))
mnd.p.16 <- as.data.frame(read_csv("./rawdata/MND-P-B-16.csv",col_names = TRUE))

mnd.roc.i <- as.data.frame(read_csv("./rawdata/MND-I-B-RoC.csv",col_names = TRUE))

mnd.roc.p <- as.data.frame(read_csv("./rawdata/MND-P-B-RoC.csv",col_names = TRUE))

# Run function and change name
mnd.incidence <- mb.gather(mnd.incidence)
colnames(mnd.incidence)[2] <- "incidence 2017"

mnd.i.12 <- mb.gather(mnd.i.12)
colnames(mnd.i.12)[2] <- "incidence 2012"
mnd.i.13 <- mb.gather(mnd.i.13)
colnames(mnd.i.13)[2] <- "incidence 2013"
mnd.i.14 <- mb.gather(mnd.i.14)
colnames(mnd.i.14)[2] <- "incidence 2014"
mnd.i.15 <- mb.gather(mnd.i.15)
colnames(mnd.i.15)[2] <- "incidence 2015"
mnd.i.16 <- mb.gather(mnd.i.16)
colnames(mnd.i.16)[2] <- "incidence 2016"
mnd.roc.i <- mb.gather(mnd.roc.i)
colnames(mnd.roc.i)[2] <- "ROC incidence"

mnd.prev <- mb.gather(mnd.prev)
colnames(mnd.prev)[2] <- "prevalance 2017"
mnd.p.12 <- mb.gather(mnd.p.12)
colnames(mnd.p.12)[2] <- "prevalance 2012"
mnd.p.13 <- mb.gather(mnd.p.13)
colnames(mnd.p.13)[2] <- "prevalance 2013"
mnd.p.14 <- mb.gather(mnd.p.14)
colnames(mnd.p.14)[2] <- "prevalance 2014"
mnd.p.15 <- mb.gather(mnd.p.15)
colnames(mnd.p.15)[2] <- "prevalance 2015"
mnd.p.16 <- mb.gather(mnd.p.16)
colnames(mnd.p.16)[2] <- "prevalance 2016"
mnd.roc.p <- mb.gather(mnd.roc.p)
colnames(mnd.roc.p)[2] <- "ROC prevalance"

# Join them together.

mnd.i <- list( mnd.i.12, mnd.i.13,mnd.i.14,mnd.i.15,mnd.i.16,mnd.incidence,mnd.roc.i) %>% reduce(left_join, by = "Location")

mnd.p <- list( mnd.p.12, mnd.p.13,mnd.p.14,mnd.p.15,mnd.p.16,mnd.incidence,mnd.roc.p) %>% reduce(left_join, by = "Location")

# For IPD

ipd.incidence <- as.data.frame(read_csv("./rawdata/PD-I-B.csv",col_names = TRUE))

ipd.i.12 <- as.data.frame(read_csv("./rawdata/PD-I-B-12.csv",col_names = TRUE))
ipd.i.13 <- as.data.frame(read_csv("./rawdata/PD-I-B-13.csv",col_names = TRUE))
ipd.i.14 <- as.data.frame(read_csv("./rawdata/PD-I-B-14.csv",col_names = TRUE))
ipd.i.15 <- as.data.frame(read_csv("./rawdata/PD-I-B-15.csv",col_names = TRUE))
ipd.i.16 <- as.data.frame(read_csv("./rawdata/PD-I-B-16.csv",col_names = TRUE))

ipd.prev <- as.data.frame(read_csv("./rawdata/PD-P-B.csv",col_names = TRUE))
ipd.p.12 <- as.data.frame(read_csv("./rawdata/PD-P-B-12.csv",col_names = TRUE))
ipd.p.13 <- as.data.frame(read_csv("./rawdata/PD-P-B-13.csv",col_names = TRUE))
ipd.p.14 <- as.data.frame(read_csv("./rawdata/PD-P-B-14.csv",col_names = TRUE))
ipd.p.15 <- as.data.frame(read_csv("./rawdata/PD-P-B-15.csv",col_names = TRUE))
ipd.p.16 <- as.data.frame(read_csv("./rawdata/PD-P-B-16.csv",col_names = TRUE))

ipd.roc.i <- as.data.frame(read_csv("./rawdata/PD-I-B-RoC.csv",col_names = TRUE))

ipd.roc.p <- as.data.frame(read_csv("./rawdata/PD-P-B-RoC.csv",col_names = TRUE))

# Run function and change name
ipd.incidence <- mb.gather(ipd.incidence)
colnames(ipd.incidence)[2] <- "incidence 2017"

ipd.i.12 <- mb.gather(ipd.i.12)
colnames(ipd.i.12)[2] <- "incidence 2012"
ipd.i.13 <- mb.gather(ipd.i.13)
colnames(ipd.i.13)[2] <- "incidence 2013"
ipd.i.14 <- mb.gather(ipd.i.14)
colnames(ipd.i.14)[2] <- "incidence 2014"
ipd.i.15 <- mb.gather(ipd.i.15)
colnames(ipd.i.15)[2] <- "incidence 2015"
ipd.i.16 <- mb.gather(ipd.i.16)
colnames(ipd.i.16)[2] <- "incidence 2016"
ipd.roc.i <- mb.gather(ipd.roc.i)
colnames(ipd.roc.i)[2] <- "ROC incidence"

ipd.prev <- mb.gather(ipd.prev)
colnames(ipd.prev)[2] <- "prevalance 2017"
ipd.p.12 <- mb.gather(ipd.p.12)
colnames(ipd.p.12)[2] <- "prevalance 2012"
ipd.p.13 <- mb.gather(ipd.p.13)
colnames(ipd.p.13)[2] <- "prevalance 2013"
ipd.p.14 <- mb.gather(ipd.p.14)
colnames(ipd.p.14)[2] <- "prevalance 2014"
ipd.p.15 <- mb.gather(ipd.p.15)
colnames(ipd.p.15)[2] <- "prevalance 2015"
ipd.p.16 <- mb.gather(ipd.p.16)
colnames(ipd.p.16)[2] <- "prevalance 2016"
ipd.roc.p <- mb.gather(ipd.roc.p)
colnames(ipd.roc.p)[2] <- "ROC prevalance"

# Join them together.

ipd.i <- list( ipd.i.12, ipd.i.13,ipd.i.14,ipd.i.15,ipd.i.16,ipd.incidence,ipd.roc.i) %>% reduce(left_join, by = "Location")

ipd.p <- list( ipd.p.12, ipd.p.13,ipd.p.14,ipd.p.15,ipd.p.16,ipd.incidence,ipd.roc.p) %>% reduce(left_join, by = "Location")

# For AD

ad.incidence <- as.data.frame(read_csv("./rawdata/AD-I-B.csv",col_names = TRUE))

ad.i.12 <- as.data.frame(read_csv("./rawdata/AD-I-B-12.csv",col_names = TRUE))
ad.i.13 <- as.data.frame(read_csv("./rawdata/AD-I-B-13.csv",col_names = TRUE))
ad.i.14 <- as.data.frame(read_csv("./rawdata/AD-I-B-14.csv",col_names = TRUE))
ad.i.15 <- as.data.frame(read_csv("./rawdata/AD-I-B-15.csv",col_names = TRUE))
ad.i.16 <- as.data.frame(read_csv("./rawdata/AD-I-B-16.csv",col_names = TRUE))

ad.prev <- as.data.frame(read_csv("./rawdata/AD-P-B.csv",col_names = TRUE))
ad.p.12 <- as.data.frame(read_csv("./rawdata/AD-P-B-12.csv",col_names = TRUE))
ad.p.13 <- as.data.frame(read_csv("./rawdata/AD-P-B-13.csv",col_names = TRUE))
ad.p.14 <- as.data.frame(read_csv("./rawdata/AD-P-B-14.csv",col_names = TRUE))
ad.p.15 <- as.data.frame(read_csv("./rawdata/AD-P-B-15.csv",col_names = TRUE))
ad.p.16 <- as.data.frame(read_csv("./rawdata/AD-P-B-16.csv",col_names = TRUE))

ad.roc.i <- as.data.frame(read_csv("./rawdata/AD-I-B-RoC.csv",col_names = TRUE))

ad.roc.p <- as.data.frame(read_csv("./rawdata/AD-P-B-RoC.csv",col_names = TRUE))

# Run function and change name
ad.incidence <- mb.gather(ad.incidence)
colnames(ad.incidence)[2] <- "incidence 2017"

ad.i.12 <- mb.gather(ad.i.12)
colnames(ad.i.12)[2] <- "incidence 2012"
ad.i.13 <- mb.gather(ad.i.13)
colnames(ad.i.13)[2] <- "incidence 2013"
ad.i.14 <- mb.gather(ad.i.14)
colnames(ad.i.14)[2] <- "incidence 2014"
ad.i.15 <- mb.gather(ad.i.15)
colnames(ad.i.15)[2] <- "incidence 2015"
ad.i.16 <- mb.gather(ad.i.16)
colnames(ad.i.16)[2] <- "incidence 2016"
ad.roc.i <- mb.gather(ad.roc.i)
colnames(ad.roc.i)[2] <- "ROC incidence"

ad.prev <- mb.gather(ad.prev)
colnames(ad.prev)[2] <- "prevalance 2017"
ad.p.12 <- mb.gather(ad.p.12)
colnames(ad.p.12)[2] <- "prevalance 2012"
ad.p.13 <- mb.gather(ad.p.13)
colnames(ad.p.13)[2] <- "prevalance 2013"
ad.p.14 <- mb.gather(ad.p.14)
colnames(ad.p.14)[2] <- "prevalance 2014"
ad.p.15 <- mb.gather(ad.p.15)
colnames(ad.p.15)[2] <- "prevalance 2015"
ad.p.16 <- mb.gather(ad.p.16)
colnames(ad.p.16)[2] <- "prevalance 2016"
ad.roc.p <- mb.gather(ad.roc.p)
colnames(ad.roc.p)[2] <- "ROC prevalance"

# Join them together.

ad.i <- list( ad.i.12, ad.i.13,ad.i.14,ad.i.15,ad.i.16,ad.incidence,ad.roc.i) %>% reduce(left_join, by = "Location")

ad.p <- list( ad.p.12, ad.p.13,ad.p.14,ad.p.15,ad.p.16,ad.incidence,ad.roc.p) %>% reduce(left_join, by = "Location")

```