### 2019-07-11 Earth_microbiome_MS.Rmd
updated

---
title: "Earth_microbiome_MS"
author: "Muriel Sirgi"
date: "03/07/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#setting the working directory in globalmicrobiome (GitHub directory)
knitr::opts_knit$set(root.dir = '~/globalmicrobiome')

```

##Data Set-up
```{r R Library Set-up}

## Get data

#libraries for data processing and formatting
	library(foreign)
	library(stargazer)
	library(psych)
	library(plyr)
	library(xtable)
	library(car) #

	#library(dplyr)

	library(scales)
	library(reshape)
	library(gmodels) #
  library(nnet)
	library(tidyr) #works well with dplyr data pipelines

	library(sjmisc)
	library(RColorBrewer)
  library(htmltools) #
	library(broom)

  #library(hrbrthemes)
library(gcookbook)
library(tidyverse)
 #library(gdtools)

#The phyloseq package seeks to address issues with multiple microbiome analysis packages by providing a set of functions that internally manage the organizing, linking, storing, and analyzing of phylogenetic sequencing data. In general, this package is used for UniFrac analyses.
library(phyloseq)
library(microbiome)
library(knitr)
library(ggpubr)
#library(phylogeo)
library(dplyr)
#Graphing package used in phyloseq. To edit the default setting of a plot, you need to use functions in this package.
library(ggplot2)
library(viridis) 
library(cowplot)
# install.packages("devtools")
library(devtools)
# install_github("opisthokonta/tsnemicrobiota")
library(tsnemicrobiota)
# check you've installed this library
library(forcats)
#package required for biomformat
library(jsonlite)
library(Matrix)
library(rhdf5)
#package used to work with biom files in R 
library(biomformat)

```

##Microbiome Data Import
```{r Microbiome Data Import}
#file.choose() has been chosen here to take away any errors pertaining to the file path while investigating the files from the ftp directory
githubdata_biom = file.choose()

#This function transforms a biom file into a phyloseq class object
githubdata_phyloseq = read_biom2phyloseq(githubdata_biom)

```

```{r Import GitHUb file (metada)}
read.table
```

```{r Microbiome Data Import Draft}
#import_biom("~/globalmicrobiome/emp_cr_silva_16S_123qc_filtered.biom") 
#import_biom(BIOMfilename = '~/globalmicrobiome/emp_cr_gg_13_8.release1.biom')
#biomfile <- file.choose("~/globalmicrobiome/emp_cr_gg_13_8.release1.biom")
#read_biom2phyloseq(biom.file = biomfile)
#import_biom('~/globalmicrobiome/emp_cr_gg_13_8.qc_filtered.biom')
#import_biom('~/globalmicrobiome/emp_cr_silva_16S_123.qc_filtered.biom'

#silva_closedref_biom <- system.file("extdata", "/Users/Muriel/globalmicrobiome/emp_cr_silva_16S_123qc_filtered.biom", package = "phyloseq")
#sample_metadata("/Users/Muriel/globalmicrobiome/emp_cr_silva_16S_123qc_filtered.biom")

#silva_closedref_biom = system.file("extdata", "~/globalmicrobiome/emp_cr_silva_16S_123qc_filtered.biom", package = "biomformat")

#silva_data = sample_metadata(silva_closedref_biom)

#biomfile = "/Users/Muriel/globalmicrobiome/emp_cr_silva_16S_123qc_filtered.biom"
#read_biom(biomfile)
#x = read_biom(biomfile)
#header(x)
#biom_shape(x)
#nrow(x)
#ncol(x)
#rownames(x)
#colnames(x)
#matrix_element_type(x)
#biom_data(x)

```

#Microbiome Functions (Nik)
```{r functions (Nik)}
##### FUNCTIONS ########
##### Normalization #######

# Better rounding function than R's base round
myround <- function(x) { trunc(x + 0.5) }

# Scales reads by 
# 1) taking proportions
# 2) multiplying by a given library size of n
# 3) rounding 
# Default for n is the minimum sample size in your library
# Default for round is floor
scale_reads <- function(physeq, n = min(sample_sums(physeq)), round = "floor") {

  # transform counts to n
  physeq.scale <- transform_sample_counts(physeq, 
    function(x) {(n * x/sum(x))}
  )

  # Pick the rounding functions
  if (round == "floor"){
    otu_table(physeq.scale) <- floor(otu_table(physeq.scale))
  } else if (round == "round"){
    otu_table(physeq.scale) <- myround(otu_table(physeq.scale))
  }

  # Prune taxa and return new phyloseq object
  physeq.scale <- prune_taxa(taxa_sums(physeq.scale) > 0, physeq.scale)
  return(physeq.scale)
}

##### ADONIS ###########

# Function to run adonis test on a phyloseq object and a variable from metadata
# Make sure OTU data is standardized/normalized before 
phyloseq_to_adonis <- function(physeq, distmat = NULL, dist = "bray", formula) {

  if(!is.null(distmat)) {
    phydist <- distmat
  } else {
    phydist <- phyloseq::distance(physeq, dist)
  }

  metadata <- as(sample_data(physeq), "data.frame")

  # Adonis test
  f <- reformulate(formula, response = "phydist")
  adonis.test <- adonis(f, data = metadata)
  print(adonis.test)

  # Run homogeneity of dispersion test if there is only 1 variable
  if (grepl("\\+", formula)) {
    l <- list(
      dist = phydist, 
      formula = f, 
      adonis = adonis.test
    )
  } else {
    group <- metadata[ ,formula]
    beta <- betadisper(phydist, group)
    disper.test = permutest(beta)
    print(disper.test)

    l <- list(
      dist = phydist, 
      formula = f, 
      adonis = adonis.test, 
      disper = disper.test
    )
  }
  return (l)
}

########## Bar Plots #################

# This function takes a phyloseq object, agglomerates OTUs to the desired taxonomic rank, 
# prunes out OTUs below a certain relative proportion in a sample (ie 1% ) 
# and melts the phyloseq object into long format which is suitable for ggplot stacked barplots.
taxglom_and_melt <- function(physeq, taxrank, prune){

  # Agglomerate all otu's by given taxonomic level
  pglom <- tax_glom(physeq, taxrank = taxrank)

  # Create a new phyloseq object which removes taxa from each sample below the prune parameter
  pglom_prune <- transform_sample_counts(pglom,function(x) {x/sum(x)})
  otu_table(pglom_prune)[otu_table(pglom_prune) < prune] <- 0
  pglom_prune <- prune_taxa(taxa_sums(pglom_prune) > 0, pglom_prune)

  # Melt into long format and sort by taxonomy
  physeq_long <- psmelt(pglom_prune)
  physeq_long <- physeq_long[order(physeq_long[ ,taxrank]), ]

  # Return long data frame
  return(physeq_long)
}

###### Merge functions ############

# Merge samples by averaging OTU counts instead of summing
merge_samples_mean <- function(physeq, group, round){
  # Calculate the number of samples in each group
  group_sums <- as.matrix(table(sample_data(physeq)[ ,group]))[,1]

  # Merge samples by summing
  merged <- merge_samples(physeq, group)

  # Divide summed OTU counts by number of samples in each group to get mean
  # Calculation is done while taxa are columns
  x <- as.matrix(otu_table(merged))
  if(taxa_are_rows(merged)){ x<-t(x) }

  # Pick the rounding functions
  if (round == "floor"){
    out <- floor(t(x/group_sums))
  } else if (round == "round"){
    out <- myround(t(x/group_sums))
  }

  # Return new phyloseq object with taxa as rows
  out <- otu_table(out, taxa_are_rows = TRUE)
  otu_table(merged) <- out
  return(merged)
}

# Merge samples, just including OTUs that were present in all merged samples
# Call this function before running merge_samples()
merge_OTU_intersect <- function(physeq, group){

  # Make sure we're not starting with more taxa than we need 
  physeq <- prune_taxa(taxa_sums(physeq) > 0, physeq)

  s <- data.frame(sample_data(physeq))
  l <- levels(s[,group])
  o <- otu_table(physeq)

  # Loop through each category
  for (cat in 1:length(l)) {

    # Get the index of all samples in that category
    w <- which(s[,group]==l[cat])

    # subset to just those columns of OTU table
    cat.sub<-o[,w]
    print(dim(cat.sub))

    # Find the indices of 0's in the OTU table
    zeros <- apply(cat.sub, 1, function(r) any(r == 0))

    # If an OTU had a 0 in at least one sample, change all samples to 0
    cat.sub[zeros,] <- 0
  }

  o[,w] <- cat.sub
  otu_table(physeq) <- o

  return(physeq)

}

```

```{r }

```