### 2019-07-15 Earth_microbiome_MS.Rmd
updated

---
title: "Earth_microbiome_MS"
author: "Muriel Sirgi"
date: "03/07/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#setting the working directory in globalmicrobiome (GitHub directory)
knitr::opts_knit$set(root.dir = '~/globalmicrobiome')

```

##Data Set-up
```{r R Library Set-up}

## Get data

#libraries for data processing and formatting
	library(foreign)
	library(stargazer)
	library(githubdata_phyloseqych)
	library(plyr)
	library(xtable)
	library(car) #

	#library(dplyr)

	library(scales)
	library(reshape)
	library(gmodels) #
  library(nnet)
	library(tidyr) #works well with dplyr data pipelines

	library(sjmisc)
	library(RColorBrewer)
  library(htmltools) #
	library(broom)

  #library(hrbrthemes)
library(gcookbook)
library(tidyverse)
 #library(gdtools)

#The phyloseq package seeks to address issues with multiple microbiome analysis packages by providing a set of functions that internally manage the organizing, linking, storing, and analyzing of phylogenetic sequencing data. In general, this package is used for UniFrac analyses.
library(phyloseq)
library(microbiome)
library(knitr)
library(ggpubr)
#library(phylogeo)
library(dplyr)
#Graphing package used in phyloseq. To edit the default setting of a plot, you need to use functions in this package.
library(ggplot2)
library(viridis) 
library(cowplot)
# install.packages("devtools")
library(devtools)
# install_github("opisthokonta/tsnemicrobiota")
library(tsnemicrobiota)
# check you've installed this library
library(forcats)
#package required for biomformat
library(jsonlite)
library(Matrix)
library(rhdf5)
#package used to work with biom files in R 
library(biomformat)

```

##Microbiome Data Import
```{r Microbiome Data Import}
#file.choose() has been chosen here to take away any errors pertaining to the file path while investigating the files from the ftp directory
githubdata_biom <- file.choose()

#This function transforms a biom file into a phyloseq class object
githubdata_phyloseq <- read_biom2phyloseq(githubdata_biom)

```

```{r Microbiome Data Import Draft}
#import_biom("~/globalmicrobiome/emp_cr_silva_16S_123qc_filtered.biom") 
#import_biom(BIOMfilename = '~/globalmicrobiome/emp_cr_gg_13_8.release1.biom')
#biomfile <- file.choose("~/globalmicrobiome/emp_cr_gg_13_8.release1.biom")
#read_biom2phyloseq(biom.file = biomfile)
#import_biom('~/globalmicrobiome/emp_cr_gg_13_8.qc_filtered.biom')
#import_biom('~/globalmicrobiome/emp_cr_silva_16S_123.qc_filtered.biom'

#silva_closedref_biom <- system.file("extdata", "/Users/Muriel/globalmicrobiome/emp_cr_silva_16S_123qc_filtered.biom", package = "phyloseq")
#sample_metadata("/Users/Muriel/globalmicrobiome/emp_cr_silva_16S_123qc_filtered.biom")

#silva_closedref_biom = system.file("extdata", "~/globalmicrobiome/emp_cr_silva_16S_123qc_filtered.biom", package = "biomformat")

#silva_data = sample_metadata(silva_closedref_biom)

#biomfile = "/Users/Muriel/globalmicrobiome/emp_cr_silva_16S_123qc_filtered.biom"
#read_biom(biomfile)
#x = read_biom(biomfile)
#header(x)
#biom_shape(x)
#nrow(x)
#ncol(x)
#rownames(x)
#colnames(x)
#matrix_element_type(x)
#biom_data(x)

#ftpdata_silva_qrbiom = import_qiime("/Users/Muriel/globalmicrobiome/emp_cr_silva_16S_123qc_filtered.biom")
#ftpdata_silva_phyloseq = read_biom2phyloseq(ftpdata_silva_qrbiom)

```

#Microbiome Functions (Nik)
```{r functions (Nik)}
##### FUNCTIONS ########
##### Normalization #######

# Better rounding function than R's base round
myround <- function(x) { trunc(x + 0.5) }

# Scales reads by 
# 1) taking proportions
# 2) multiplying by a given library size of n
# 3) rounding 
# Default for n is the minimum sample size in your library
# Default for round is floor
scale_reads <- function(physeq, n = min(sample_sums(physeq)), round = "floor") {

  # transform counts to n
  physeq.scale <- transform_sample_counts(physeq, 
    function(x) {(n * x/sum(x))}
  )

  # Pick the rounding functions
  if (round == "floor"){
    otu_table(physeq.scale) <- floor(otu_table(physeq.scale))
  } else if (round == "round"){
    otu_table(physeq.scale) <- myround(otu_table(physeq.scale))
  }

  # Prune taxa and return new phyloseq object
  physeq.scale <- prune_taxa(taxa_sums(physeq.scale) > 0, physeq.scale)
  return(physeq.scale)
}

##### ADONIS ###########

# Function to run adonis test on a phyloseq object and a variable from metadata
# Make sure OTU data is standardized/normalized before 
phyloseq_to_adonis <- function(physeq, distmat = NULL, dist = "bray", formula) {

  if(!is.null(distmat)) {
    phydist <- distmat
  } else {
    phydist <- phyloseq::distance(physeq, dist)
  }

  metadata <- as(sample_data(physeq), "data.frame")

  # Adonis test
  f <- reformulate(formula, response = "phydist")
  adonis.test <- adonis(f, data = metadata)
  print(adonis.test)

  # Run homogeneity of dispersion test if there is only 1 variable
  if (grepl("\\+", formula)) {
    l <- list(
      dist = phydist, 
      formula = f, 
      adonis = adonis.test
    )
  } else {
    group <- metadata[ ,formula]
    beta <- betadisper(phydist, group)
    disper.test = permutest(beta)
    print(disper.test)

    l <- list(
      dist = phydist, 
      formula = f, 
      adonis = adonis.test, 
      disper = disper.test
    )
  }
  return (l)
}

########## Bar Plots #################

# This function takes a phyloseq object, agglomerates OTUs to the desired taxonomic rank, 
# prunes out OTUs below a certain relative proportion in a sample (ie 1% ) 
# and melts the phyloseq object into long format which is suitable for ggplot stacked barplots.
taxglom_and_melt <- function(physeq, taxrank, prune){

  # Agglomerate all otu's by given taxonomic level
  pglom <- tax_glom(physeq, taxrank = taxrank)

  # Create a new phyloseq object which removes taxa from each sample below the prune parameter
  pglom_prune <- transform_sample_counts(pglom,function(x) {x/sum(x)})
  otu_table(pglom_prune)[otu_table(pglom_prune) < prune] <- 0
  pglom_prune <- prune_taxa(taxa_sums(pglom_prune) > 0, pglom_prune)

  # Melt into long format and sort by taxonomy
  physeq_long <- psmelt(pglom_prune)
  physeq_long <- physeq_long[order(physeq_long[ ,taxrank]), ]

  # Return long data frame
  return(physeq_long)
}

###### Merge functions ############

# Merge samples by averaging OTU counts instead of summing
merge_samples_mean <- function(physeq, group, round){
  # Calculate the number of samples in each group
  group_sums <- as.matrix(table(sample_data(physeq)[ ,group]))[,1]

  # Merge samples by summing
  merged <- merge_samples(physeq, group)

  # Divide summed OTU counts by number of samples in each group to get mean
  # Calculation is done while taxa are columns
  x <- as.matrix(otu_table(merged))
  if(taxa_are_rows(merged)){ x<-t(x) }

  # Pick the rounding functions
  if (round == "floor"){
    out <- floor(t(x/group_sums))
  } else if (round == "round"){
    out <- myround(t(x/group_sums))
  }

  # Return new phyloseq object with taxa as rows
  out <- otu_table(out, taxa_are_rows = TRUE)
  otu_table(merged) <- out
  return(merged)
}

# Merge samples, just including OTUs that were present in all merged samples
# Call this function before running merge_samples()
merge_OTU_intersect <- function(physeq, group){

  # Make sure we're not starting with more taxa than we need 
  physeq <- prune_taxa(taxa_sums(physeq) > 0, physeq)

  s <- data.frame(sample_data(physeq))
  l <- levels(s[,group])
  o <- otu_table(physeq)

  # Loop through each category
  for (cat in 1:length(l)) {

    # Get the index of all samples in that category
    w <- which(s[,group]==l[cat])

    # subset to just those columns of OTU table
    cat.sub<-o[,w]
    print(dim(cat.sub))

    # Find the indices of 0's in the OTU table
    zeros <- apply(cat.sub, 1, function(r) any(r == 0))

    # If an OTU had a 0 in at least one sample, change all samples to 0
    cat.sub[zeros,] <- 0
  }

  o[,w] <- cat.sub
  otu_table(physeq) <- o

  return(physeq)

}

```

#Read phyloseq object
```{r Phyloseq object - first read}
#Methods from MRD microbiome file
#Print phyloseq object to check its component parts
print(githubdata_phyloseq)

#Check sparsity of physeq data
kable(head(tax_table(githubdata_phyloseq)))
tax_table(githubdata_phyloseq)

#summarize_phyloseq(githubdata_phyloseq)
#The above code ran unsuccessfully due to the fact that the data does not include sample data
```

#Taxonomic filtering
```{r Taxonomic filtering}
#Methods obtained from Callahan et al.

#Show taxonomic ranks in the phyloseq object
rank_names(githubdata_phyloseq)

#Create table showing the phyla present in the table 
table(tax_table(githubdata_phyloseq)[, "Phylum"], exclude = NULL)
#It is noted here that some phyla are named following a different convention

#Remove ambiguous phyla (if any)
githubdata_phyloseq <- subset_taxa(githubdata_phyloseq, !is.na(Phylum) & !Phylum %in% c("", "uncharacterized"))

#Compute prevalence of each Phyla, store as data.frame
prevdf = apply(X = otu_table(githubdata_phyloseq),
               MARGIN = ifelse(taxa_are_rows(githubdata_phyloseq), yes = 1, no = 2),
               FUN = function(x){sum(x > 0)})

#Add taxonomy and total read counts to this data.frame
prevdf = data.frame(Prevalence = prevdf,
                    TotalAbundance = taxa_sums(githubdata_phyloseq),
                    tax_table(githubdata_phyloseq))

#Create table dispaying the average prevalence of each phyla per sample and the total reads for all phyla in the data frame
plyr::ddply(prevdf, "Phylum", function(df1){cbind(mean(df1$Prevalence),sum(df1$Prevalence))})

```

#READ ME - Problems encountered

Files downloaded from the emp ftp directory could not be reecognised a JSON or HDF5 file as required by the biomformat package and therefore could not be read in R. This is a significant problem that explains the use of a GitHub file (emp_deblur_150bp.subset_2k.rare_5000.biom").
This file was used as a test to verify the working order of certain sections of the code.

Table selected to run code ("/Users/Muriel/Downloads/emp_deblur_150bp.subset_2k.rare_5000.biom") does not feature sample data - it may be necessary to look into other files or acquire the sample data from another file using the sample references

Some of the phyla in the table are named following a differenct convention. If this file is used for the data this will have to be kept in mind. It is possible that other files for EMP data also share this problem.