### 2019-08-09 IPCO_NS.Rmd
updated

---
title: "Untitled"
author: "NSharma"
date: "5 May 2016"
output: html_document
---

```{r setup, include=FALSE}

	library(foreign)
	library(stargazer)
	library(psych)
	library(plyr)
	library(xtable)
	library(car)
	#library(dplyr)
	library(scales)
	library(reshape) 
	library(gmodels)
	library(nnet)
	library(tidyr)
	library(sjmisc)
	library(RColorBrewer)
  library(htmltools)
	library(broom)

library(gcookbook)
library(tidyverse)
 #library(gdtools)

#The phyloseq package seeks to address issues with multiple microbiome analysis packages by providing a set of functions that internally manage the organizing, linking, storing, and analyzing of phylogenetic sequencing data. In general, this package is used for UniFrac analyses.
library(phyloseq)
library(microbiome)
library(knitr)
library(ggpubr)
#library(phylogeo)
library(dplyr)
#Graphing package used in phyloseq. To edit the default setting of a plot, you need to use functions in this package.
library(ggplot2)
library(viridis) 
library(cowplot)
# install.packages("devtools")
library(devtools)
# install_github("opisthokonta/tsnemicrobiota")
library(tsnemicrobiota)
# check you've installed this library
library(forcats)
library(IPCO)
library(phylosmith)
library(phyloseq)

data(atlas1006)   
pseq <- baseline(atlas1006)

##Step 1: Extract the required matrix from Phyloseq object
##first convert the OTU level phyloseq object to species/genus level phyloseq object. Depending on the phyloseq version the taxrank or taxlevel parameter would be used.
#taxlevel parameter is probably for older version to the best of my knowledge

#phyloseq_genus <- tax_glom(pseq, taxlevel="Genus")

#taxrank parameter is in the latest version
phyloseq_genus <- tax_glom(pseq, taxrank="Genus")

#If the above step is already done, then you can start from here.
#Convert phyloseq genus/species level object to matrix
genus_table <- as(otu_table(phyloseq_genus), "matrix")

##Step 2: Using IPCO
#load refernce data IPCO_Healthy/IPCO_HMP
library(IPCO)
data(IPCO_Healthy)

#Do normalisation/transform as suggested as an where required. If the phyloseq data is already normalised and doesnot wish to renormalise, then the user doesnot have to follow IPCO normalisation.
#We recommend Hellinger transformation. To be used on count/prop/relative abundance data
ref_genus <- transform_data(IPCO_Healthy$Genus,"relab")
ref_func <- transform_data(IPCO_Healthy$MetaCyc,"prop")
##or
ref_func <- transform_data(IPCO_Healthy$KEGG,"prop")

#if query genus table is count data
query_genus <- transform_data(genus_table,"count")

#Run IPCO
#Note: all samples should be in columns and features (func/taxa) in rows. If not, transpose the matrix
inferred_MetaCyc <- IPCO(ref_func,ref_genus,query_genus)

data <- data("IPCO_Healthy")

write_phyloseq(pseq, type = "all", path = getwd())

tdata <- transform_data(data, type = "proportion")
tdata <- transform_data(IPCO_Healthy, proportion)

```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.