### 2019-08-12 IPCO_NS.Rmd
updated

---
title: "Untitled"
author: "NSharma"
date: "2019"
output: html_document
---

```{r setup, include=FALSE}

library(gcookbook)
library(tidyverse)
 #library(gdtools)

#The phyloseq package seeks to address issues with multiple microbiome analysis packages by providing a set of functions that internally manage the organizing, linking, storing, and analyzing of phylogenetic sequencing data. In general, this package is used for UniFrac analyses.
library(phyloseq)
library(microbiome)
library(knitr)
library(ggpubr)
#library(phylogeo)
library(dplyr)
#Graphing package used in phyloseq. To edit the default setting of a plot, you need to use functions in this package.
library(ggplot2)
library(viridis) 
library(cowplot)
# install.packages("devtools")
library(devtools)
# install_github("opisthokonta/tsnemicrobiota")
library(tsnemicrobiota)
# check you've installed this library
library(forcats)
library(IPCO)
library(phylosmith)
library(phyloseq)

data(atlas1006)   
pseq <- baseline(atlas1006)

##Step 1: Extract the required matrix from Phyloseq object
##first convert the OTU level phyloseq object to species/genus level phyloseq object. Depending on the phyloseq version the taxrank or taxlevel parameter would be used.
#taxlevel parameter is probably for older version to the best of my knowledge

#phyloseq_genus <- tax_glom(pseq, taxlevel="Genus")

#taxrank parameter is in the latest version
phyloseq_genus <- tax_glom(pseq, taxrank="Genus")

#If the above step is already done, then you can start from here.
#Convert phyloseq genus/species level object to matrix
genus_table <- as(otu_table(phyloseq_genus), "matrix")

##Step 2: Using IPCO
#load refernce data IPCO_Healthy/IPCO_HMP
library(IPCO)
data(IPCO_Healthy)

#Do normalisation/transform as suggested as an where required. If the phyloseq data is already normalised and doesnot wish to renormalise, then the user doesnot have to follow IPCO normalisation.

#We recommend Hellinger transformation. To be used on count/prop/relative abundance data
ref_genus <- transform_data(IPCO_Healthy$Genus,"relab")
ref_func <- transform_data(IPCO_Healthy$MetaCyc,"proportion")

#ref_func <- filter_functionality(ref_func, "metacyc",IPCO_Healthy$MetaCyc_coverage,threshold=0.4) # Filtering

ref_func <- filter_functionality(ref_func, "metacyc", IPCO_Healthy$MetaCyc_coverage)
##or
#ref_func <- transform_data(IPCO_Healthy$KEGG,"proportion")

#if query genus table is count data
#query_genus <- transform_data(genus_table,"count")

#Run IPCO
#Note: all samples should be in columns and features (func/taxa) in rows. If not, transpose the matrix
inferred_MetaCyc <- IPCO(ref_func,ref_genus,genus_table)

library(made4)

overview(inferred_MetaCyc)

k.coa<- ord(inferred_MetaCyc, type="coa")
summary(k.coa$ord)

plotgenes(k.coa)

heatmap(inferred_MetaCyc)

p <- plot_ly(inferred_MetaCyc = volcano, type = "heatmap")

levelplot(inferred_MetaCyc[1:ncol(inferred_MetaCyc),ncol(inferred_MetaCyc):1])

coinertia_resuls <- check_coinertia(ref_func,genus_table)

```

## R Markdown