### 2019-10-15 curatedMetagenomic_NS.Rmd
updated

---
title: "curatedMetagenomicData"
author: "NSharma"
date: "26/09/2019"
output: html_document
---

```{r setup, include=FALSE}

#BiocManager::install("curatedMetagenomicData")
#BiocManager::install("ExperimentHub")
library(ExperimentHub)
library(curatedMetagenomicData)
suppressPackageStartupMessages(library(ExperimentHub))
library(phyloseq)
library(microbiome)
library(forcats)
library(stargazer)
library(ggpubr)
library(tidyverse)
library(cowplot)
library(vegan)
library(broom)
library(tsnemicrobiota)
library(latex2exp)

eh = ExperimentHub()
myquery = query(eh, "curatedMetagenomicData")

# Get the stool samples
esl <- curatedMetagenomicData("*metaphlan_bugs_list.stool*", dryrun = FALSE)

# merge
eset <- mergeData(esl)

# the relab=FALSE is essential as it Absolute Raw Count Data
pseq = ExpressionSet2phyloseq(eset,relab=FALSE ,simplify = TRUE, phylogenetictree = TRUE)
orig_pseq <- pseq
#For the MetaPhlAn2 bugs datasets (but not other data types), you gain a lot of taxonomy-aware, ecological analysis and plotting by conversion to a phyloseq class object. curatedMetagenomicData provides the ExpressionSet2phyloseq() function to make this easy:

```

## R Markdown

```{r}
# Create mnd variable
pseq <- orig_pseq
# grabs the nationality from phyloseq
mnd.country <- get_variable(pseq, "country" )
table(mnd.country)

# this collapses the two varialbes into a new variable AB in this case. The command for multiple changes is fct_collapse(x, AB = c("A","B"), DE = c("D","E"))
# Not used BGD Bangladesh , CAN Canada, CHN, China. DNK denmark, FJI Figi, HUN Hungry, ISL isle of man , MDG, MNG mongolia, PER Iran, RUS RUSSIA,

mnd.country <- fct_collapse(mnd.country, LOW = c("NOR","SWE","FIN","DEN","LUX") , MEDIUM = c("ITA","AUT","DEU","ESP","FRA","NLD"), HIGH = c("GBR","USA","CAN") )

#mnd.country <- fct_collapse(mnd.country, LOW = c("Scandinavia","EasternEurope") , MEDIUM = c("SouthEurope","CentralEurope", "UKIE"), HIGH = c("US") )

# reorder
mnd.country <- factor(mnd.country, levels = (c("LOW", "MEDIUM", "HIGH")))
levels(mnd.country)

# creates a new variable in the phyloseq called mnd
sample_data(pseq)$mnd = mnd.country

# checks that it has worked.
get_variable(pseq, "mnd")

```

## Check OTUs/Phylums
Prune(remove) taxa because we want to remove OTUs/taxa that aren't in any of the samples
Don't need to prune this taxa because OTUs are found in all samples.
Check this everytime you use the subset_samples function because of the pseq object being filtered. 
```{r check OTUs}
# Check if any OTUs are not present in any samples
any(taxa_sums(pseq) == 0)
# FALSE, therefore, OTUs are found in all samples and we don't need to prune the taxa. If the answer is TRUE, some OTUs are not present in any samples. This is usually the case when data is subset to remove some samples. OTUs unique to those sample are not removed along with the samples. Therefore, it is important to check this everytime the phyloseq object is filtered for samples using `subset_samples` function.

head(meta(pseq))

```

## Prepare Data for Alpha Diversity Visualisation
```{r prepare data for alpha diversity visualisation}

# https://bioconductor.riken.jp/packages/3.5/data/experiment/vignettes/curatedMetagenomicData/inst/doc/curatedMetagenomicData.html
# Prepare data for vizualisation

# Remove   everything without a country

keepotu = genefilter_sample(pseq, filterfun_sample(topp(0.05)), A=5)
pseq.f <- subset_taxa(pseq, keepotu)

pseq.c <- core(pseq, detection = 0.1/100, prevalence = 50/100)

pseq.subset <- subset_samples(pseq.c, !is.na(mnd) & disease == "healthy" & age_category != "child" & age_category != "newborn" & age_category != "schoolage" & antibiotics_current_use != "yes")

pseq.subset <- subset_samples(pseq.c, !is.na(mnd) & disease == "healthy" & age_category != "child" & age_category != "newborn" & age_category != "schoolage" )

# These are useful is filtering. 
#sample_variables(pseq.subset)
#get_variable(pseq.subset, sample_variables(pseq.subset)[13])

any(taxa_sums(pseq.subset) == 0)

# run the diversities. 	keep an eye on this as only one is working
tab <- diversities(pseq.subset, index = "Shannon")

# Get metadata
pseq.meta <- meta(pseq.subset)
#kable(head(pseq.meta))
stargazer(pseq.meta, type = "text",summary = NULL)

# THis subsets the  genus and phylum. 
#genus.pseq <- subset_taxa( pseq.subset, !is.na(Genus))
#phylum.pseq <- subset_taxa( pseq, is.na(Class) & !is.na(Phylum))

# Shannon
# the number in the palettes must equal the number of groups in bmi.pairs
p1 <- ggviolin(data=subset(pseq.meta, !is.na(mnd)), x = "mnd", y = "Shannon", add = "boxplot", fill = "mnd", na.rm = TRUE, palette = c("yellow", "orange", "red"))
p1 <- p1 + stat_compare_means(comparisons = mnd.pairs) + xlab("MND Prevalence") + theme(legend.position="none")
print(p1)
ggsave("mnd_violin__shannon_stats.png")

alpha_d <- function(pseq.subset) {
   # gets the meta data 
   pseq.meta <- meta(pseq.subset)
   # generates the diversites
   tab <- diversities(pseq.subset, index = "Shannon")
   # creates Shannon
   pseq.meta$Shannon <- tab$shannon
   mnd <- levels(pseq.meta$mnd) # get the variables
  # Make a pairwise list that we want to compare.
  mnd.pairs <- combn(seq_along(mnd), 2, simplify = FALSE, FUN = function(i)mnd[i])
  mnd.pairs <- list( c("LOW","MEDIUM"), c("HIGH","LOW"), c("MEDIUM","HIGH"))

   # Compare differences in Shannon index between mnd group of the study subjects
  pseq.meta$mnd <- factor(pseq.meta$mnd, levels = c("LOW","MEDIUM","HIGH"),  exclude = "NA")
  # on to the graphs!!!
  p1 <-  ggviolin(pseq.meta, x = "mnd", y = "Shannon",
 add = "boxplot", fill = "mnd", palette = "paired") + theme(axis.text.x = element_text(angle = 90)) + theme(legend.position="none") + xlab("")
  p2<-   ggviolin(pseq.meta, x = "country", y = "Shannon",
 add = "boxplot", fill = "country", palette = "paired") + theme(axis.text.x = element_text(angle = 90)) + theme(legend.position="none") + xlab("")
# the number in the palettes must equal the number of groups in bmi.pairs
  p3 <- ggviolin(data=subset(pseq.meta, !is.na(mnd)), x = "mnd", y = "Shannon", add = "boxplot", fill = "mnd", na.rm = TRUE, palette = c("yellow", "orange", "red"))
  p3 <- p3 + stat_compare_means(comparisons = mnd.pairs) + xlab("MND Prevalence") + theme(legend.position="none")
  print(p1)
  print(p2)
  p4 <-plot_grid(p1, p2,p3, align = "h" , ncol = 3)

}

beta_d <- function(pseq.subset) {
  p.low <- subset_samples(pseq.subset, mnd == "LOW")
  p.low.d <- divergence(p.low)
  p.med <- subset_samples(pseq.subset, mnd == "MEDIUM")
  p.med.d <- divergence(p.med)
  p.high <- subset_samples(pseq.subset, mnd == "HIGH")
  p.high.d <- divergence(p.high)
  p <- boxplot(list(Low = p.low.d, Medium = p.med.d, High = p.high.d))
  #plot(p)
  # NMDS on the whole set???
  pseq.ord <- ordinate(pseq.subset, "NMDS", "bray")
  # Convert to compositional data
  pseq.rel <- microbiome::transform(pseq.subset, "compositional")
  otu <- abundances(pseq.rel)
  meta <- meta(pseq.rel)
  pseq.nmds <- ordinate(pseq.rel, "NMDS", "bray")
  #NMDS ordination on Bray-Curtis distance
  p1 = plot_ordination(pseq.subset, pseq.nmds, type = "taxa", color = "mnd", title = "Taxa by Phylum")
  p1 + facet_wrap(~Phylum, 3)
  p2 = plot_ordination(pseq.rel, pseq.nmds, type = "sample", color = "mnd")
  plot(p1 + scale_colour_manual(values = c("yellow", "orange", "red", "grey")) + labs(color = "MND Prevalence"))
  p2 + scale_colour_manual(values = c("yellow", "orange", "red", "grey")) + facet_wrap(~mnd, nrow = 2) + labs(color = "MND Prevalence")

    # MDS
   pseq.mds <- ordinate(pseq.rel, "MDS", "bray")
   p3 = plot_ordination(pseq.rel, pseq.mds, type = "sample", color = "mnd")
 plot(p1 + scale_colour_manual(values = c( "yellow", "orange", "red", "grey")) + labs(color = "MND Prevalence"))

   p3 + scale_colour_manual(values = c("yellow", "orange", "red", "grey")) + facet_wrap(~mnd, 3) + labs(color = "MND Prevalence")
   #PCoA for compositional data with Bray-Curtis distances
    #   p4 <- plot_landscape(pseq.rel, method = "NMDS", distance = "bray", colour = "mnd", size = 3) 
    #   p4 <- p4 + scale_color_brewer(palette = "Dark2")+ scale_fill_gradient(low = "#e0ecf4", high = "#6e016b") 
      p5 <- plot_landscape(pseq.rel, method = "MDS", distance = "bray", col = "mnd", size = 3) 
  print(p5)
  # 
  #   # Pick core taxa
    pseq.core <- core(pseq.rel, detection = 5/100, prevalence = 5/100)
  # 
  #   # PCoA for compositional data with Bray-Curtis distances
   p6 <- plot_landscape(microbiome::transform(pseq.core, "compositional"),
                        method = "PCoA", distance = "bray") +
         labs(title = paste("PCoA / Compositional / Bray-Curtis"))
  # 
    tsne_res <- tsne_phyloseq(pseq.subset, distance='bray', perplexity = 5, verbose=1, rng_seed = 3901)
    p7 <- plot_tsne_phyloseq(pseq.subset, tsne_res,
    color = 'mnd')
    p7 + scale_colour_manual(values = c( "yellow", "orange", "red", "grey")) + labs(color = "MND Prevalence")

  print(p7)

}

mylab <- function(v1) {
  deparse(substitute(v1))
}

pre_perma_d <- function(pseq.subset) { 
  p.low <<- subset_samples(pseq.subset, mnd == "LOW")
  print(bob)
  p.low.d <<- divergence(p.low)
  p.med <<- subset_samples(pseq.subset, mnd == "MEDIUM")
  p.med.d <<- divergence(p.med)
  p.high <<- subset_samples(pseq.subset, mnd == "HIGH")
  p.high.d <<- divergence(p.high)
  p <<- boxplot(list(Low = p.low.d, Medium = p.med.d, High = p.high.d))

}

pre_perma_d(pseq.subset) 
perma_d(p.low,p.high)

perma_d <- function(x,y) { # x & y must be either p.low,p.med or p.high

  # Convert to compositional data
  pseqLvH <- merge_phyloseq(x,y)
  tempx <- mylab(x)
  pseq.rel <- microbiome::transform(pseqLvH, "compositional")
  otu <- abundances(pseq.rel)
  meta <- meta(pseq.rel)
  permanova <- adonis(t(otu) ~ mnd,
               data = meta, permutations=999, method = "bray")
 # P-value
  print(as.data.frame(permanova$aov.tab)["mnd", "Pr(>F)"])
  dist <- vegdist(t(otu))
  anov <- anova(betadisper(dist, meta$mnd))
  summary(anov)
  tidy(anov)
  coef <- coefficients(permanova)["mnd1",]
  top.coef <- coef[rev(order(abs(coef)))[1:20]]
  par(mar = c(3, 14, 2, 1))
  temp <- TeX("Top Genus $tempx")
  barplot(sort(top.coef), col = rainbow(20), horiz = T, las = 1, main = temp)
 }

## MICROBIOME COMPOSITION
paste("Hello", "world", sep=" ")

comp_d <- function(pseq,mylabel) {
  png(paste(mylabel,"Phylum_abundance.png",sep="_"))
  p <- plot_bar(pseq,"Phylum") + theme(axis.text.x = element_text(face="italic"))
  print(p)

  prev.otu <- plot_taxa_prevalence(pseq, "Phylum")
  prev.otu + theme(legend.position="none")
  ggsave(paste(mylabel,"phylum_prevalence.png",sep="_"))

  tab <- plot_taxa_prevalence(pseq, "Phylum", detection = 5)
  tab + theme(legend.position="none")
  ggsave("Phylum_prevalence_detection5.png")

  pseq.ph <- aggregate_taxa(pseq, "Phylum")
  pseq.ph.rel <- microbiome::transform(pseq.ph, "compositional")
  plot.comp.rel <- plot_composition(pseq.ph.rel, x.label = "sample") +
    theme(legend.position = "bottom") + theme_bw() +
    theme(axis.text.x=element_blank()) +
    c + scale_fill_brewer(palette = "Paired") + guides(fill=guide_legend("Phylum"))
  print(plot.comp.rel)
  ggsave("phylum_relative_abundance11.png")
  plot.comp.rel + theme(legend.position="none")
  ggsave("phylum_relative_abundance12.png")

  pseq.ph <- aggregate_taxa(pseq, "Phylum")
  pseq.ph.rel <- microbiome::transform(pseq.ph, "compositional")
  plot.comp.rel <- plot_composition(pseq.ph.rel, average_by = "mnd", x.label = "MND   Incidence", plot.type = "barplot") + guides(fill=guide_legend(title="Phylum"))
  plot.comp.rel + scale_fill_viridis(discrete = TRUE, option = "plasma") 
  ggsave("phylum_abundance_bar.png")
  # Top 5
  top5P.names = sort(tapply(taxa_sums(pseq), tax_table(pseq)[, "Genus"], sum), TRUE)[1:5]

  top5P = subset_taxa(pseq, Genus %in% names(top5P.names))
  pseq.ph.rel <- microbiome::transform(top5P, "compositional")
  plot.comp.rel <- plot_composition(pseq.ph.rel, average_by = "mnd", x.label = "MND Incidence", plot.type = "barplot") + guides(fill=guide_legend(title="Phylum"))

  plot.comp.rel + scale_fill_viridis(discrete = TRUE, option = "plasma")

  # Top 10
  top10P.names = sort(tapply(taxa_sums(pseq), tax_table(pseq)[, "Genus"], sum), TRUE)[1:10]

  top10P = subset_taxa(pseq, Genus %in% names(top10P.names))
  pseq.ph.rel <- microbiome::transform(top10P, "compositional")

  plot.comp.rel <- plot_composition(pseq.ph.rel, average_by = "mnd", x.label = "MND Incidence", plot.type = "barplot") + guides(fill=guide_legend(title="Phylum"))

  plot.comp.rel + scale_fill_viridis(discrete = TRUE, option = "plasma") 
  # new
  p.new <- aggregate_taxa(pseq, "Phylum")
  p.new.rel <- microbiome::transform(p.new, "compositional")

  plot.p.new.rel <- plot_composition(p.new.rel, average_by = "mnd", x.label = " ", plot.type = "barplot") + guides(fill=guide_legend(title="Phylum")) + theme(legend.text = element_text( face="italic"))

  plot.p.new.rel + scale_fill_brewer(palette = "Paired") 
  ggsave("pnew_abundance_bar.png")

  # Top Phylum Heatmap
  pseq.ph <- aggregate_taxa(pseq, "Phylum")

  pseq.ph.rel <- microbiome::transform(pseq.ph, "compositional")

  plot.comp.rel <- plot_composition(pseq.ph.rel, x.label = "sample", plot.type = "heatmap")

  plot(plot.comp.rel+ coord_flip()) + theme(axis.text.x=element_blank())
  ggsave("phylum_histogram.png")

  #p.new heatmap
  pseq.n.ph <- aggregate_taxa(pseq, "Phylum")
  pseq.n.ph.rel <- microbiome::transform(pseq.n.ph, "compositional")

  plot.n.comp.rel <- plot_composition(pseq.n.ph.rel, x.label = "sample", plot.type = "heatmap")

  p <- plot.n.comp.rel + coord_flip()
  p + theme(axis.text.x=element_blank()) + labs(y = "Sample")
  ggsave("pnew_histogram.png")

  # Newwork
  png("network_bray_phylum.png")
  jg = make_network(pseq, "taxa", "bray", 0.3)
  plot_network(jg, pseq, "taxa", color = "Phylum", line_weight = 0.4, label = NULL)
  dev.off()
  #Community-composition-plots
  # Create a data table for ggplot
# agglomerate at phylum level
pseq_phy <- tax_glom(pseq.new, taxrank = "Phylum")

# Transform to rel. abundance (or use ps0.ra)
pseq_phyl <- transform_sample_counts(pseq_phy, function(x) {x/sum(x)})

# Melt to long format for easy ggploting
pseq_phylum <- psmelt(pseq_phyl)

     # filter(Abundance > 0.01)                             # Filter out low abundance taxa
# Plot - Phylum
  p.ra.phylum <- ggplot(pseq_phylum, aes(x = sample, y = Abundance, fill = Phylum)) + 
  geom_bar(stat = "identity", width = 1) +
  facet_wrap(~mnd, scales = "free_x", nrow = 4, ncol = 7) +
  theme(axis.text.x = element_blank()) +
  theme(axis.title.x = element_blank()) +
  labs(title = "Abundant Phylum (> 1%)")

  p <- p.ra.phylum + scale_fill_viridis(discrete = TRUE, option = "A")
  plot(p)
  ggsave("phylum_abun_mnd.png")
}

alpha_d(pseq.subset         )
beta_d(pseq.subset)

```

```{r MND table summary}

library(table1)
n <- length(sample_data(pseq.subset)$mnd)

data <- data.frame(id=1:n)
data$mnd <- sample_data(pseq.subset)$mnd
data$bmi <- sample_data(pseq.subset)$BMI
data$age <- sample_data(pseq.subset)$age
data$antibiotics_current_use <- sample_data(pseq.subset)$antibiotics_current_use
data$smoker <- sample_data(pseq.subset)$smoker
data$pregnant <- sample_data(pseq.subset)$pregnant
data$alcohol <- sample_data(pseq.subset)$alcohol
data$age_category <- sample_data(pseq.subset)$age_category

table1(~ age  +  bmi + smoker + alcohol + age_category | mnd, data=data)

```

## BETA DIVERSITY
Beta-diversity: Measures for differences between samples from different groups to identify if there are differences in the overall community composition and structure.
WE want to: quantify community divergence within a given sample set
Divergence of a given sample can be quantified as the average dissimilarity of each sample from the group mean; the dissimilarity can be quantified by beta diversity.

Divergence of a given sample set can be quantified as the average dissimilarity of each sample from the group mean; the dissimilarity can be quantified by beta diversity, for instance. This was applied in group-level comparisons for instance in Salonen et al. ISME J 2014 (they focused on homogeneity using inverse correlation, whereas here we focus on divergence using correlation but the measure is essentially the same).

 The low group tends to have smaller values, indicating that the samples are more similar to the group mean, and is less heterogeneous (has smaller spread / is more homogeneous):

## PERMANOVA significance test for group-level differences
Evaluate whether the MND grouping has a significant effect on overall gut microbiota composition

## MICROBIOME COMPOSITION

```{r Phylum #}

# Show available ranks in the dataset
rank_names(pseq)

# Create table, number of features for each phyla
table(tax_table(pseq)[, "Phylum"], exclude = NULL)

```
This shows a few phyla for which only one feature was observed. Those may be worth filtering.

guide_italics
ggtitle("Relative abundance") +  
```{r}

clos.ph <- aggregate_taxa(clostridium, "Phylum")

clos.ph.rel <- microbiome::transform(clos.ph, "compositional")

plot.comp.rel <- plot_composition(clos.ph.rel, x.label = "sample") + 
  theme(legend.position = "bottom", ) + theme_bw() + 
  theme(axis.text.x=element_blank()) + 
  theme(legend.title = element_text(size=18), legend.text = element_text( face="italic")) + scale_fill_brewer(palette = "Paired") + guides(fill=guide_legend(title="Phylum"))

print(plot.comp.rel)

ggsave("phylum_relative_abundance_clos.png")

plot.comp.rel + scale_fill_brewer(palette = "Paired") + theme(legend.position="none") + guide_italics
ggsave("phylum_relative_abundance_clos2.png")

```

```{r CLostridium Bar Plot}

clos <- aggregate_taxa(clostridium, "Phylum")

pseq.cl.rel <- microbiome::transform(clos, "compositional")

plot.cl.rel <- plot_composition(pseq.cl.rel, average_by = "mnd", x.label = "sample", plot.type = "barplot") + guides(fill=guide_legend(title="Phylum")) + theme(legend.text = element_text( face="italic"))

plot.cl.rel + scale_fill_brewer(palette = "Paired")
ggsave("cl_abundance_bar.png")

```

```{r Clostridium genus bar plots}

pseq.cl.rel <- microbiome::transform(clostridium, "compositional")

plot.cl.rel <- plot_composition(pseq.cl.rel, average_by = "mnd", x.label = "sample", plot.type = "barplot")

plot.cl.rel + scale_fill_viridis(discrete = TRUE, option = "A")

```

```{r clostridium heatmap}

pseq.c.ph <- aggregate_taxa(clostridium, "Phylum")

pseq.c.ph.rel <- microbiome::transform(pseq.c.ph, "compositional")

plot.c.comp.rel <- plot_composition(pseq.c.ph.rel, x.label = "sample", plot.type = "heatmap")

p <- plot.c.comp.rel + coord_flip()
p + theme(axis.text.x=element_blank()) + labs(y = "Sample")
ggsave("clos_histogram.png")

```

## Community composition plotting
Classic bar plots of bacterial phyla present per sample can be useful for communicating "high level" results. These are relatively easy to interpret when major shifts in microbial communities are present, such as in this study where antibiotics are used However, they are not effective at detecting subtle shifts in communities or taxa and do not convey any statistical significance and can be subjectively interpreted. Interpretation of these plots should always be subject to subsequent statistical analysis.

```{r}

clos <- aggregate_taxa(clostridium, "Phylum")

# agglomerate at phylum level
clos_phy <- tax_glom(clos, taxrank = "Phylum")

# Transform to rel. abundance (or use ps0.ra)
clos_phyl <- transform_sample_counts(clos_phy, function(x) {x/sum(x)})

# Melt to long format for easy ggploting
clos_phylum <- psmelt(clos_phyl)

# Plot - Phylum
p.ra.phylum <- ggplot(clos_phylum, aes(x = sample, y = Abundance, fill = Phylum)) + 
  geom_bar(stat = "identity", width = 1) +
  facet_wrap(~mnd, scales = "free_x", nrow = 4, ncol = 7) +
  theme(axis.text.x = element_blank()) +
  theme(axis.title.x = element_blank()) +
  labs(title = "Abundant Phylum of Clostridium(> 1%)")

p <- p.ra.phylum + scale_fill_viridis(discrete = TRUE, option = "A")
plot(p)
ggsave("phylum_abun_mnd_clos.png")

```

```{r}

get_taxa_unique(pseq, "Phylum")

```

```{r}

get_taxa_unique(pseq, "Genus")

```

```{r}

get_taxa_unique(clostridium, "Genus")

```

```{r}

df <- psmelt(pseq)
kable(head(df))

```

```{r}

head(abundances(pseq)["Faecalibacterium prausnitzii et rel.",])

```